---
title: "p04_eval_model"
author: "okura"
date: "2023-11-10"
output: html_document
editor_options: 
  chunk_output_type: console
---

# load package
```{r}
require(tidyverse)
require(lubridate)
require(survival)
require(nlme)
require(JM)
require(openxlsx)
require(officer)
require(survivalROC)
require(ggpubr)
```

# definition; ggsave2_tri, sv2ggkm
```{r}

outputdir <- "./figs/"
if(!dir.exists(outputdir)){
  dir.create(outputdir)
}

ggsave2_tri <- function(plot, wid = 9, hei=9){
  plot_name <- deparse(substitute(plot))
  
  #png
  file_name <- paste(outputdir,plot_name, ".png", sep = "",collapse = "")
  ggsave(filename = file_name,plot = plot,device = "png",width = wid, height = hei,dpi = 300,units = "cm")
  
  
  #pdf
  file_name <- paste(outputdir,plot_name, ".pdf", sep = "",collapse = "")
  
  #pptx
  file_name <- paste(outputdir,plot_name, ".pptx", sep = "",collapse = "")
  plot_ed <- rvg::dml(ggobj = plot,width = wid, height = hei)
  
  doc <- read_pptx() %>%
    add_slide('Title and Content', 'Office Theme') %>%
    ph_with(value = "", location = ph_location_type(type = "title")) %>%
    ph_with(value = plot_ed, location = ph_location(type = "body",width = cm2inch(wid), height = cm2inch(hei),left = 2, top = 2))
  doc %>%
    print(target = file_name)
}

inch2cm <- function(x){
  return(x/0.39370)
}
cm2inch <- function(x){
  return(0.39370*x)
}

p2cm <- function(x){
  return(as.integer(0.008466666666666667 * x))
}

cm2p <- function(x){
  return(as.integer(118.11023622047244 *x))
}

sv2ggkm <- function(svfit, df){
  p <- ggsurvplot(fit = svfit, data = df,
                risk.table.title = "risk.table.title", risk.table.y.text.col = TRUE,
           risk.table.y.text = FALSE, fontsize = 5,
           risk.table = TRUE, tables.theme = theme_pubr(), ggtheme = theme_pubr())
  p_km <- p$plot
  p_tb <- p$table
  p2 <- cowplot::plot_grid(p_km, p_tb, ncol = 1,align = "v", rel_heights = c(3,1))
  grid::grid.draw(p2)
  return(p2)
}
```

# definition; get_survfitJM_mst
```{r}
get_survfitJM_mst <-
  function (x, estimator = c("both", "mean", "median"), 
            which = NULL, fun = NULL, conf.int = FALSE,
            fill.area = FALSE, col.area = "grey", col.abline = "black", col.points = "black",
            add.last.time.axis.tick = FALSE, include.y = FALSE, main = NULL, xlab = NULL, ylab = NULL, ylab2 = NULL,
            lty = NULL, col = NULL, lwd = NULL, pch = NULL, ask = NULL, legend = FALSE, ..., cex.axis.z = 1, 
        cex.lab.z = 1) {
    #fix from plot.survfitJM
    estimator <- match.arg(estimator)
    fun <- if (!is.null(fun)) match.fun(fun)
    if (is.null(which))
        which <- seq_along(x$summaries)
    if (conf.int && is.null(x$success.rate)) {
        warning("a confidence interval can be included only when argument", 
            "'simulate' of survfitJM() was set to TRUE.")
        conf.int <- FALSE
    }
    if (is.null(ask))
        ask <- prod(par("mfcol")) < length(which)
    if (ask) {
        op <- par(ask = TRUE)
        on.exit(par(op))
    }
    if (is.null(main)) {
        main <- paste("Subject", names(x$summaries))
        names(main) <- names(x$summaries)
    }
    if (is.null(xlab))
        xlab <- rep("Time", length(which))
    if (is.null(ylab)) {
        ylab <- if (is.null(fun))
            if (!include.y)
                rep(expression(paste("Pr(", T[i] >= u, " | ", T[i] > t, 
                    ", ", tilde(y)[i](t), ")", sep = " ")), length(which))
            else
                rep("Survival Probability", length(which))
        else
            rep("", length(which))
    }
    if (is.null(ylab2))
        ylab2 <- "Longitudinal Outcome"
    if (!is.null(x$success.rate)) {
        if (is.null(col))
            col <- switch(estimator, both = c(2, 3, 1, 1), mean = c(2, 1, 1), median = c(3, 1, 1))
        if (is.null(lty))
            lty <- switch(estimator, both = c(1, 1, 2, 2), mean = c(1, 2, 2), median = c(1, 2, 2))
        if (is.null(lwd))
            lwd <- switch(estimator, both = c(1, 1, 1, 1), mean = c(1, 1, 1), median = c(1, 1, 1))
    } else {
        col <- lty <- lwd <- 1
    }
    if (is.null(pch))
        pch <- 8
    l_eval <- list()
    for (i in seq_along(which)) {
        ii <- which[i]
        r <- x$summaries[[ii]]
        r <- if (!is.null(x$success.rate)) {
            rbind(cbind(c(0, x$last.time[ii]), matrix(1, 2, 4)), r)
        } else {
            rbind(cbind(c(0, x$last.time[ii]), matrix(1, 2, 1)), r)
        }
        if (!is.null(fun) && is.function(fun))
            r[, 2:ncol(r)] <- fun(r[, 2:ncol(r)])
        if (!is.null(x$success.rate) && estimator == "mean")
            r <- r[, -3]
        if (!is.null(x$success.rate) && estimator == "median")
            r <- r[, -2]
        if (!conf.int && !is.null(x$success.rate)) {
            exc <- c(ncol(r) - 1, ncol(r))
            r <- r[, -exc, drop = FALSE]
            col <- col[-exc]
            lty <- lty[-exc]
            lwd <- lwd[-exc]
        }
        ylim <- if (is.null(fun)) c(0, 1) else { rr <- r[, -1, drop = FALSE]; range(rr[is.finite(rr)]) } 
        
        lt <- x$last.time[ii]
        r. <- r[r[, 1] >= lt, ]
        l_eval <- append(l_eval,list(r.))
        rng <- range(x$obs.times[[ii]], x$survTimes)
    
    }
    # invisible()
    
    mst <- get_mst(l_eval)
    mst <- mst -as.numeric(x$last.time)
    return(mst)
    # return(l_eval)
}

get_mst <- function(l_eval){
  df_xx<- l_eval %>% 
    as.data.frame() 
  if(0.5 < min(df_xx$Mean)){
    return(NA)
  }else{
    
  df_xx <- df_xx%>% 
    as.data.frame() %>% 
    mutate(y = Mean -0.5) %>% 
    mutate(ud = y>0) %>% 
    group_by(ud) %>% 
    mutate(min_y = min(abs(y))) %>% 
    filter(min_y == abs(y))
  x1 <- df_xx$times[[1]]
  x2 <- df_xx$times[[2]]
  y1 <- df_xx$y[[1]]
  y2 <- df_xx$y[[2]]
  a <- (y2 -y1)/(x2 - x1)
  b <- a*x1 -y1
  x0 <- x1 -(b/a)*y1
  return(x0)
  }
}
```


# load and save functions
```{r}
append_list <- function(l, x){
  xname  <- deparse(substitute(x))
  l[[xname]] <- x
  return(l)
}

# load function
get_vals <- function(l){
  lnames <- names(l)
  for(n in lnames){
    assign(n, l[[n]], pos = .GlobalEnv)
  }
  return(lnames)
}
```

# timestamp
```{r}
ts <- Sys.time()  %>% 
    as.character() %>% 
    str_remove_all(., "[-:]*") %>% 
    str_replace(.," ", "_")
ts
```

# load files
```{r}
f_rds <- "../p03_modeling/dat/eval_list_20240702_083021.206594.rds"
readRDS(f_rds) %>%
  get_vals()
```

# measured CEA vs predicted CEA(mixed-effects model)
```{r}
df_mix_predict <- df_mix_cea %>%
    mutate(xx = predict(fit_mix_cea)) %>% 
    mutate(yy = log2(CEA))

mix_prediction <- ggplot()+theme_pubr(base_family = "Times")+
  theme(aspect.ratio = 1)+
  geom_point(aes(x = xx, y = yy), data=df_mix_predict, size=0.01) +
  xlab("Predicted log2(CEA)")+ylab("Measured log2(CEA)")

plot(mix_prediction)

ggsave2_tri(mix_prediction)

cor.test(df_mix_predict$xx, df_mix_predict$yy, method = "pearson")
```

# MSE 
```{r}
df_mix_predict_sub <- df_mix_predict %>% 
  mutate(error_2 = (xx-yy)^2) %>% 
  summarise(mse = mean(error_2))

df_mix_predict_sub
```

# Clinical Case Presentation.1
```{r}
options(repr.plot.width=6, repr.plot.height=6)

fig_pred_83 <- ggplot() + theme_pubr(base_size = 15, base_family = "Times") +
  geom_line(aes(x = year_samp_from_start, y = xx, color = id),
                data = df_mix_predict[df_mix_predict$id %in% c(83:83), ], 
                color = "red")+
  geom_point(aes(x = year_samp_from_start, y = log2(CEA), color = id),
                data = df_mix_predict[df_mix_predict$id %in% c(83:83), ],
                size = 0.8, color = "black")+
  coord_cartesian(xlim = c(0, 8),ylim = c(-1,4), expand = T) + 
  xlab("years") + ylab("log2(CEA)")
        
plot(fig_pred_83)

ggsave2_tri(fig_pred_83)
```

# Clinical Case Presentation.1-survival probability
```{r}
d_83 <- df_mix_predict %>% 
  filter(id ==83)

#time = 5, 10, 15, 20
par(mfrow = c(2, 2))
for (i in c(5, 10, 15, 20)) {
  sfit83_i <- survfitJM(fit_JM_cea, newdata = d_83[1:i, ], idVar = "id")
  fig83<- plot(sfit83_i, estimator="mean", include.y = TRUE, conf.int=0.95, fill.area=TRUE,
       col.area="lightblue", main="Patient 83", ask = F)
}
```

# Clinical Case Presentation.1-MST
```{r}
for (i in c(5, 10, 15, 20)) {
  sfit83_i <- survfitJM(fit_JM_cea, newdata = d_83[1:i, ], idVar = "id")

  fig_r <- c("fig_83_", i, ".png") %>%
    str_c(collapse = "")

  mst<- get_survfitJM_mst(sfit83_i, estimator="mean", include.y = TRUE,
                          conf.int=0.95, fill.area=TRUE,col.area="lightblue",
                          main="Patient 83", ask = F)
  print(i)
  print(mst)
}
```

# Clinical Case Presentation.2
```{r}
options(repr.plot.width=6, repr.plot.height=6)

fig_pred_317 <- ggplot() + theme_pubr(base_size = 15, base_family = "Times") +
  geom_line(aes(x = year_samp_from_start, y = xx, color = id),
                data = df_mix_predict[df_mix_predict$id %in% c(317:317), ], 
                color = "red")+
  geom_point(aes(x = year_samp_from_start, y = log2(CEA), color = id),
                data = df_mix_predict[df_mix_predict$id %in% c(317:317), ],
                size = 0.8, color = "black")+
  coord_cartesian(xlim = c(0, 10),ylim = c(0,8), expand = T) + 
  xlab("years") + ylab("log2(CEA)")
        
plot(fig_pred_317)

ggsave2_tri(fig_pred_317, wid = 9, hei=9)
```

# Clinical Case Presentation.2-survival probability
```{r}
d_317 <- df_mix_predict %>% 
  filter(id == 317)

#time = 25, 50, 75, 100
par(mfrow = c(2, 2))
fig_317 <- for (i in c(25, 50, 75, 100) ) {
  sfit317_i <- survfitJM(fit_JM_cea, newdata = d_317[1:i, ], idVar = "id")
  plot317_i <- plot(sfit317_i, estimator="mean", include.y = TRUE, conf.int=0.95, fill.area=TRUE,
       col.area="lightblue", main="Patient 317", ask = F)
}
```

# Clinical Case Presentation.2-MST
```{r}
for (i in c(25, 50, 75, 100)) {
  sfit317_i <- survfitJM(fit_JM_cea, newdata = d_317[1:i, ], idVar = "id")

  fig_r <- c("fig_317_", i, ".png") %>%
    str_c(collapse = "")

  mst<- get_survfitJM_mst(sfit317_i, estimator="mean", include.y = TRUE,
                          conf.int=0.95, fill.area=TRUE,col.area="lightblue",
                          main="Patient 317", ask = F)
  print(i)
  print(mst)
}
```


# Figure_whole
```{r}
df_whole <- df_mix_cea %>% 
  dplyr::select(id, CEA, year_samp_from_start) %>% 
  rename(x = year_samp_from_start) %>% 
  rename(y = CEA) %>% 
  mutate(y = log2(y)) %>% 
  mutate(id = as.numeric(id))

fig_whole <- ggplot() + theme(legend.position = "none") + theme_pubr(base_size = 20, base_family = "Times") + 
              labs(y = "Measured" ~ log[2]("CEA")) + scale_x_continuous(name = "(years)") 

for(i in c(1:max(df_whole$id))){
  fig_whole <- fig_whole +
               geom_line(data = df_whole[df_whole$id ==i, ], aes(x = x, y= y), color = "black", linetype = "solid", linewidth = 0.3)
}

fig_whole <- fig_whole + 
  geom_smooth(data = df_whole, aes(x = x, y = y), method = "lm", se = FALSE, color = "red") 

fig_whole

ggsave2_tri(fig_whole)
```

# time-dependent ROC:each model
```{r}
ROC_cox_sub_fix <- function(cutoff, cox_model){
  
  x_cox <- predict(cox_model, type = "lp")
  y_time <- cox_model$y[,1]
  y_event <- cox_model$y[,2]

  span_ <- 0.25 * length(y_time)^(-0.2)
  
  cox_roc <- survivalROC(Stime = y_time, status = y_event, 
                        marker = x_cox,
                        predict.time = cutoff, method = "NNE", span = span_)

  plot(x = cox_roc$FP, y = cox_roc$TP, type = "l", 
  xlim = c(0, 1), ylim = c(0, 1), 
  xlab = paste("FP", "\n", "AUC =", round(cox_roc$AUC, 3)),
  ylab = "TP", main = paste("COX-CEA ,Method = NNE \n Time =", cutoff, "(year)"))
  abline(0, 1)

  cox_roc$AUC %>% 
    return()
}

ROC_jm_sub_fix <- function(cutoff, model){
    df_jm_pred <- ranef(model) %>% 
      as.data.frame() %>% 
      `colnames<-`(c("r_intercept", "r_slope")) %>%
      mutate(id = rownames(.))
    
    df_mix_cea_sub <- df_mix_cea %>% 
      left_join(df_jm_pred, by = "id") %>% 
      mutate(prd_fix = predict(model, newdata = ., idVar = "id")) %>% 
      mutate(prd_rnd = year_samp_from_start*r_slope + r_intercept) %>% 
      mutate(CEA_pred = prd_fix + prd_rnd) %>% 
      group_by(id) %>% 
      mutate(max_year_samp_from_start_upto_cutoff = max(year_samp_from_start[year_samp_from_start <= cutoff])) %>% 
      filter(year_samp_from_start == max_year_samp_from_start_upto_cutoff)
    
    span_ <- 0.25 * (nrow(df_mix_cea_sub)^(-0.2))
    
    
    jm_roc <- survivalROC(Stime = df_mix_cea_sub$time_y, status = df_mix_cea_sub$event, 
                          marker = df_mix_cea_sub$CEA_pred, predict.time = cutoff,
                          method = "NNE", span = span_)
    
    fig_jm <- plot(x = jm_roc$FP, y = jm_roc$TP, type = "l", 
        xlim = c(0, 1), ylim = c(0, 1),
        xlab = paste("FP", "\n", "AUC =", round(jm_roc$AUC, 3)),
        ylab = "TP", main = paste("JM-CEA ,Method = NNE \n Time =", cutoff, "(year)"))
        abline(0, 1)
    jm_roc$AUC %>% 
      return()
}

eval_years <- c(1, 2, 3, 4, 5)
cox_auc <- c()
jm_auc <- c()

par(mfrow = c(2, 3))
for(i in eval_years){
  cox_auc <- c(cox_auc,
               ROC_cox_sub_fix(i, fit_cox))
}

par(mfrow = c(2, 3))
for(i in eval_years){
  jm_auc <- c(jm_auc,
              ROC_jm_sub_fix(i, fit_JM_cea))
}
```

# time-dependent ROC:comparison
```{r}
df_auc <- tibble(years = eval_years, cox_auc, jm_auc) %>% 
  print() %>% 
  pivot_longer(cols = c("cox_auc", "jm_auc"), 
               names_to = "model", 
               values_to = "auc") 

fig_auc <- ggplot(df_auc, aes(x = years, y = auc, color = model)) +
  theme_pubr(base_family = "Times")+
  theme(
  axis.text.x = element_text(size = 15),
  axis.text.y = element_text(size = 15),
  axis.title = element_text(size = 25),
  legend.text = element_text(size = 20),
  legend.title = element_text(size = 20)
 )+
    geom_line(linewidth = 1) +
  geom_point(size = 2) +
  ylim(0.68, 0.88)

grid::grid.draw(fig_auc)

ggsave2_tri(fig_auc, wid = 12, hei = 9)
```

```{r}
sessionInfo()
```
